 Austral Spring Extratropical Southern Hemisphere zonally assymetric circulation

### 2021-07-26 analysis/paper/shceof.Rmd
Arregla ecuación en precipitación
(#54)

EOFs are more flexible than Fourier descomposition modes in the sense that they can characterise \textcolor{red}{Esto hay que exresarlo más claramente}.
Nonetheless, they are restricted to standing oscillation modes and could not represent properly propagating or phase-varying modes such as zonal waves.
Besides, a single EOF can also represent a mixture of two or more physical modes.
The inability of a single EOF to describe travelling waves is what forces the PSA pattern to be described by two EOFs.
@irving2016 characterises the Pacific--South American Pattern in a Fourier framework.
Being a meridionally propagating mode, the PSA cannot be correctly characterised by Fourier decomposition at each latitude circle, so they reprojected meridional wind fields so that the path of the PSA laid on the equator.
On this new projection, they could identify the PSA using Fourier decomposition.

@@

  regr_psi %>%
    .[cEOF == which_cEOF] %>%
    .[lev == which_lev] %>%
    # .[abs(estimate) > 1, estimate := NA] %>%
    .[lat %between% c(-90, 10)] %>%
    .[, p_val := Pvaluate(estimate, std.error, df, "fdr"), by = .(cEOF, term)] %>%
    .[, var := "Streamfunction"] %>%
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +
    geom_contour_fill(aes(z = estimate*1e-7, fill = ..level..),
                      breaks = AnchorBreaks(exclude = 0)) +
    geom_contour_pval(aes(z = p_val)) +
    geom_streamline(aes(dx = fx, dy = fy), min.L = 3, L = 10,
                    size = 0.1, res = 2, nx = 22, ny = 12, 
                    arrow.angle = 14, arrow.length = 0.2) +
    geom_qmap(keep = .015) +
    scale_fill_divergent_discretised("Streamfunction (*1e-7)", 
                                     guide = guide_colorstrip_bottom(20, order = 1)) +
    ggnewscale::new_scale_fill() +

@@

(ref:australia-pp-phase-cap) Same as Figure \@ref(fig:enso-phase) but for precipitation anomalies averaged between `r knitr::combine_words(LonLabel(box_australia$lon))` and `r knitr::combine_words(LatLabel(box_australia$lat))` (box shown in Figure \@ref(fig:pp-oceania) column b). 
Black line is the fit precipitation ~ sin(phase) + cos(phase) computed by weighted OLS using the magnitude of the cEOF2 as weights.

@@

equation <- latex2exp::TeX(paste0("$\\mathrm{Precipitation} = ", 
                                  signif(coef(model)[1], 2), "\\frac{mm}{day} \\sin{(\\mathrm{phase})} + ", 
                                  signif(coef(model)[2], 2), "\\frac{mm}{day} \\cos{(\\mathrm{phase})}$"), output = "text")

@@

\textcolor{red}{La formula en la figura trata de escribirla en una forma más precisa "precipitación"/"phase" son muy generales, y también teniendo en cuenta las unidades. También incluilo en el caption.}

@@

The second complex EOF represents a wave-3 pattern with maximum magnitude in the Pacific sector.
Essentially, it is an alternate representation of the PSA1 and PSA2 patterns [@mo2001].
We show that the Imaginary cEOF2 can be identified with the PSA1 and the Real cEOF2 with the PSA2.
This cEOF prefers a phase aligned with the Imaginary cEOF2.
This is because the when ENSO is active, it forces the cEOF2 to be in the Imaginary phase.
When ENSO is neutral, the cEOF2 is still active, but with no particular preferred phase.
This mirrors the results of @cai2002, who showed that the CSIRO Model can create PSA-like variability even in the absence of ENSO forcing (with a climatological run), but the variability of one of the PSA modes was enhanced when adding the ENSO signal. The magnitude of the cEOF2 appears not to be related to the magnitude of ENSO, except frot the three strongest ENSO events in the period (`r knitr::combine_words(year(gdata[order(-oni)][1:3]$time))`) which coincide with the three strongest cEOF2 years.

>>>>

EOFs are more flexible than Fourier descomposition modes in the sense that they can capture oscillation patterns that cannot be characterised by purely sinusoidal waves with constant amplitude.
Nonetheless, they are restricted to standing oscillation modes and could not represent properly propagating or phase-varying modes such as zonal waves.
A single EOF can also represent a mixture of two or more physical modes.
The inability of a single EOF to describe travelling waves is what forces the PSA pattern to be described by two EOFs.
@irving2016 characterises the Pacific--South American Pattern in a Fourier framework.
Being a meridionally propagating mode, the PSA cannot be correctly characterised by Fourier decomposition at each latitude circle, so they reprojected meridional wind fields so that the path of the PSA laid on the equator.
On this new projection, they could identify the PSA using Fourier decomposition.

@@

  regr_psi %>%
    .[cEOF == which_cEOF] %>%
    .[lev == which_lev] %>%
    # .[abs(estimate) > 1, estimate := NA] %>%
    .[lat %between% c(-90, 10)] %>%
    .[, p_val := Pvaluate(estimate, std.error, df, "fdr"), by = .(cEOF, term)] %>%
    .[, var := "Streamfunction"] %>%
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +
    geom_contour_fill(aes(z = estimate*1e-7, fill = ..level..),
                      breaks = AnchorBreaks(exclude = 0)) +
    geom_contour_pval(aes(z = p_val)) +
    geom_streamline(aes(dx = fx, dy = fy), min.L = 3, L = 10,
                    size = 0.1, res = 2, nx = 22, ny = 12, 
                    arrow.angle = 14, arrow.length = 0.2) +
    geom_qmap(keep = .015) +
    scale_fill_divergent_discretised("Streamfunction (×1e-7)", 
                                     guide = guide_colorstrip_bottom(20, order = 1)) +
    ggnewscale::new_scale_fill() +

@@

(ref:australia-pp-phase-cap) Same as Figure \@ref(fig:enso-phase) but for precipitation anomalies averaged between `r knitr::combine_words(LonLabel(box_australia$lon))` and `r knitr::combine_words(LatLabel(box_australia$lat))` (box shown in Figure \@ref(fig:pp-oceania) column b) in $\frac{\mathrm{mm}}{\mathrm{day}}$. 
Black line is the fit precipitation ~ sin(phase) + cos(phase) computed by weighted OLS using the magnitude of the cEOF2 as weights.

@@

equation <- latex2exp::TeX(paste0("$\\mathrm{Precipitation}_{\\mathrm{West Australia}} = ", 
                                  signif(coef(model)[1], 2), "\\sin{(\\mathrm{phase})} + ", 
                                  signif(coef(model)[2], 2), "\\cos{(\\mathrm{phase})}$"), 
                           output = "text")

@@

@@

The second complex EOF represents a wave-3 pattern with maximum magnitude in the Pacific sector.
Essentially, it is an alternate representation of the PSA1 and PSA2 patterns [@mo2001].
We show that the Imaginary cEOF2 can be identified with the PSA1 and the Real cEOF2 with the PSA2.
This cEOF prefers a phase aligned with the Imaginary cEOF2.
This is because the when ENSO is active, it forces the cEOF2 to be in the Imaginary phase.
When ENSO is neutral, the cEOF2 is still active, but with no particular preferred phase.
This mirrors the results of @cai2002, who showed that the CSIRO Model can create PSA-like variability even in the absence of ENSO forcing (with a climatological run), but the variability of one of the PSA modes was enhanced when adding the ENSO signal. The magnitude of the cEOF2 appears not to be related to the magnitude of ENSO, except frot the three strongest ENSO events in the period which coincide with the three strongest cEOF2 years.
